---
title: "Using gradient descent for linear regression in R"
author: "Mike Fang"
date: "March 9, 2016"
output: html_document
---
Dr. Ng showed how to use gradient descent to find the linear regression fit in matlab. Here is a demonstration of how to implement it in R. The syntax of matlab and R differs a lot in vector/matrix indexing, but the idea is the same.

First, let's generate some data to work with, let's define our training example with 4 features:
```{r Data}
set.seed(11)
x <- matrix(rnorm(400), ncol = 4)
y <- rnorm(100)
m <- length(y)
X<-cbind(rep(1, 100), x)
theta<-rep(0,5)
```
Second, set up the cost function for least square linear regression:
```{r cost}
compCost<-function(X, y, theta){
   m <- length(y)
   J <- sum((X%*%theta- y)^2)/(2*m)
return(J)
}
```
Next, set up the gradient descent function, running for iterations:
```{r grad descent}
gradDescent<-function(X, y, theta, alpha, num_iters){
       m <- length(y)
  J_hist <- rep(0, num_iters)
  for(i in 1:num_iters){

# this is a vectorized form for the gradient of the cost function
# X is a 100x5 matrix, theta is a 5x1 column vector, y is a 100x1 column vector
# X transpose is a 5x100 matrix. So t(X)%*%(X%*%theta - y) is a 5x1 column vector
    theta <- theta - alpha*(1/m)*(t(X)%*%(X%*%theta - y))
    
# this for-loop records the cost history for every iterative move of the gradient descent,
# and it is obtained for plotting number of iterations against cost history.
    J_hist[i]  <- compCost(X, y, theta)
  }
# for a R function to return two values, we need to use a list to store them:
  results<-list(theta, J_hist)
  return(results)
}
```
Then, let's set a training rate alpha and number of iterations to perform gradient descent:
```{r iteration, echo=TRUE}
alpha <- .1
num_iters <- 150
results <- gradDescent(X, y, theta, alpha, num_iters)
theta <- results[[1]]
cost_hist <- results[[2]]
print(theta)
```
Finally let's plot the cost history:
```{r plot J_history, echo = TRUE}

plot(1:num_iters, cost_hist, type = 'l')

```
